---
title: "Report on Large Spending by South Glos Council"
author: "Joseph Osborne"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report will investigate the spending patterns of South Gloucestershire Council using the data provided by said council [here](https://www.southglos.gov.uk/business/tenders-and-contracts/how-we-spend-our-money/council-payments-over-500/). 

I will use this data to answer the following questions:

1. Who are the most significant recipients of council payments?
1. Does the volume of outgoings change over time?
1. Is the majority of the total outgoing value made up of small payments?


## Data Sourcing

This report leverages R to scrape the above webpage for the data links and brings all the data together to support any conclusions. The code is abstracted for the purposes of clarity, but can be viewed and reused at leisure [here](http://a.link.webpage/).



## - TESTING SECTION 

```{r Obtain the links}
source(here::here("R/extract_functions.R"))

links <- generateLinksFromPage("https://www.southglos.gov.uk/business/tenders-and-contracts/how-we-spend-our-money/council-payments-over-500/")

```

```{r Extracting data from page}

extract <- links %>% 
        dataExtraction(control_links = FALSE)

```

```{r Munging and refining the extracted data}

refined_extract <- extract %>% 
        refineData()
        
```

```{r Generating a basic view of the spending}
library(dplyr)
date_level_spending_data <- refined_extract %>% 
        group_by(Payment.date) %>% 
        summarise(Total.Spend = sum(GL.Code.Net.Amount))



```

```{r experimenting with the plot}
library(ggplot2)
date_level_spending_plot <- date_level_spending_data %>% 
        ggplot(aes(x = Payment.date, y = Total.Spend)) +
        geom_line(colour = '#CC2D2D',size = 2,linetype = 1,alpha = 0.67) + 
    theme_classic() +
    theme(
        plot.subtitle = element_text(size = 13, hjust = 0.01, vjust = 1)
        , plot.caption = element_text(vjust = 1)#
        , axis.ticks = element_line(colour = "black")
        , axis.title = element_text(face = "bold")
        , axis.text = element_text(face = "bold")
        , plot.title = element_text(face = "bold")
        ) + 
    labs(
        title = "Total Daily Spend"
        , x = "Payment Day"
        , y = "Total Spend"
        , subtitle = "The total council spend across all recipient categories"
        , caption = "Source: South Gloucestershire Council"
        )
```


