---
title: "Report on Large Spending by South Glos Council"
author: "Joseph Osborne"
date: "`r Sys.Date()`"
output: 
        html_document:
                toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
        echo = TRUE
        , out.width = '100%'
        )
library(drake)
```

## Introduction

This report will investigate the spending patterns of South Gloucestershire Council using the data provided by said council [here](https://www.southglos.gov.uk/business/tenders-and-contracts/how-we-spend-our-money/council-payments-over-500/). 

I will use this data to answer the following questions:

1. Who are the most significant recipients of council payments?
1. Does the volume of outgoings change over time?
1. Is the majority of the total outgoing value made up of small payments?


## Data Sourcing

This report leverages R to scrape the above webpage for the data links and brings all the data together to support any conclusions. The code is abstracted for the purposes of clarity, but can be viewed and reused at leisure [here](http://a.link.webpage/).

One comment to be made is that while there seems to be a standard across open data providers to make the data available in monthly or quarterly chunks, those chunks are themselves rarely uniform. In the case of this data, there appears to be a minimal or inconsistent standard in naming files. In addition, the column names change between data which is ostensibly recording the same information. In some cases, there are files which have been saved as .xlsx format and which have additional empty sheets which serve no purpose. These files have then simply been renamed to .csv format. 

There are cases where the csv files have additional columns made up of blank entries. This tricks the reader in to looking for data in those columns, and when the extra blank rows disappear it causes parsing errors. In certain files there is also breaks between each of the Department records to include what is most likely a summary value.




```{r experimenting with the plot}


readd(date_level_spending_plot)
```

As part of the process for generating this plot, we needed to handle some improper dates. In a few cases the dates were recorded with years which could not be possible, generally either 1900 or 2892. Either a file formatting or a human error, these dates were readjusted using the filenames to be for the correct year. 

```{r Looking at spend by department}
readd(dept_level_spending_plot)
```

When looking into the spending by department, it is clearly a story of reorganisation. In 2015 and 2013 there are breaks in the departments, the inference being that departments are reorganised, with their responsibilities and therefor spending, changing at each time. In some cases this is a dramatic shift, and there are some clear stories to look into. 

1. The formation of the Children Adults and Health department, followed by 1 year later it being the largest department by spending. This could be indicative of a centralisation of responsibilities from lower levels of government or smaller departments, it could also represent a decentralisation of responsibility from higher levels of government. 
1. The sharp doubling of the yearly spending of the corporate and central services  
1. The abrupt peak in spending on Environment and Community services

Due to the somewhat confusing nature of the spending structure with respect to the department, it may be necessary to split this section into multiple ones, with more incisive charts. This could also be rectified with greater contextual information from the council itself. One way to obtain this will be to look a the council minutes across the periods of greatest change. 

## Which Creditors recieve the most money 

Its first important to note that the creditors are not consitently referred to. Normally they are a company, however sometimes they refer to an address, or even an individual.
```{r}
readd(creditor_plot)

```

Unsurprisingly, the most fundamental features of local government feature heavily. Policing, Waste Disposal, Mass Transit and Fire Services. These are also the most centralised of the government services, this becomes apparent when you realise that the Social care company Milestones Trust is in the top 20 3 times. When you combine just those 3 then that Trust has received more than the Avon Fire authority. 


## Spending on Voluntary Bodies

There is a description field in the payments data. In some cases this field reads 'Grants to Voluntary Bodies' this can be for a variety of purposes, but my hypothesis was that as the composition of the council's responsibilities changed, they may be forced to rely on charities more. An alternative is that as their budget was put under more and more pressure, that the grants to voluntary bodies would be supressed. 

```{r}
readd(voluntary_plot)
```

It turns out that neither are the case. While the amount paid in any given month fluctuates wildly, the month on month trend is flat. This indicates that there is no, or minimal, consistent change in the amounts paid in grants to voluntary bodies. 


## Sources

Below is a list of sources consulted in the generation of the report. Not all the sources contributed any unique content, but they have all been kept for their value to future investigations. 

* [A discussion around how local government is funded](https://citizensassembly.co.uk/how-is-local-government-funded/)
* [An analysis on the funding gaps felt by local government](https://www.local.gov.uk/sites/default/files/documents/Technical Document 2020.pdf)
* [A detailed discussion on business rates and government grants](https://www.local.gov.uk/topics/finance-and-business-rates/local-taxation-council-tax-and-business-rates)
* [Hansard records of relevant debates](https://hansard.parliament.uk/search/Debates?house=commons)
* [Census and government data for south gloucestershire](https://www.southglos.gov.uk/council-and-democracy/census/)
* [Town and parish level population statistics](https://www.southglos.gov.uk/documents/research/key statistics for town and parish councils.pdf)